---
title: "Lab 0"
author: "Your Name"
date: today
format: 
  pdf:
      toc: true
      toc-depth: 4
      shift-heading-level-by: 2
      fig-pos: "H"
      fig-cap-location: top
      geometry:
        - top=1in
        - right=.8in
        - bottom=1in
        - left=.8in
      link-citations: true
      linkcolor: blue
      include-in-header: 
        text: |
          \usepackage{fancyhdr}
          \usepackage{titling}
          \pagestyle{fancy}
          \fancyhf{}
          \renewcommand\maketitle{
            \fancyhead[C]{
              \thetitle
              \ifx \theauthor\empty  \else \ – \theauthor \fi
              \ifx \thedate\empty  \else \ – \thedate \ \fi
            }
          }
          \fancyfoot[C]{\thepage}
---

Everything within the section above, demarcated by three dashes (---) is called a YAML header. We provide code for each lab containing a pre-formatted header, which will specify settings for the output format of your labs to look consistent and professional. We only ask that you update the document author to your name.

## Before we start

If you have zero to no experience with R, please take a look at this tutorial: [SICSS Boot Camp (Beta)](https://sicss.io/boot_camp/)

At least do

1.  Installing R and R Studio (14 minutes)

2.  R Basics (17 minutes)

If you are comfortable navigating R and are familiar with quarto, then feel free to skip lab 0.

If you have never coded in your life, take a deep breath and don't worry! We will provide all the necessary resources for you to succeed - with just one expectation: **read the code**.

## Intro to lab format and all other stuff

What is [qmd](https://quarto.org/docs/computations/r.html), and and how to use it.

Please run the following line in your *Terminal* (not the console!)

> quarto install tinytex

## Formatting text: pretty things and code chunks

### This is a 3 level heading

# This is a 1 level heading, a.k.a. title

this is text

```{r this is an R chunk}
1 + 1
```

::: {.callout-note}
## This is the title of the callout note

This is a callout note. If you decide to resubmit a lab, please indicate any changes using this callout format.
:::

```{r}
## this is a comment
```

```{r this is a chunk}
print("I like networks 🌐")
```

## Lets save some characters into variables

```{r}
my_name <- "Megan"
# my_name <- "Seva"

print(my_name)
```

## how about some libraries (packages)

```{r}
#| output: false
#| message: false

#those comments will hide system messages from the console

# install.packages("tidyverse")
# install.packages("palmerpenguins")

library(tidyverse)
library(palmerpenguins)
```

## a bit of data science

```{r}
data(package = 'palmerpenguins')
head(penguins)

penguins |> 
  count(species)
```

```{r output=FALSE}
library(tidygraph)
library(igraph)
library(ggraph)
```

```{r}

peng_island <- penguins |> 
  count(species, island) 

peng_island |> 
  graph_from_data_frame() |> 
  as_tbl_graph() |> 
  ggraph(layout = "tree")+
  geom_edge_link(aes(alpha = n))+
  geom_node_label(aes(label = name))+
  theme_void()+
  ggtitle("Islands and penguins")
```

## let's save our data

```{r}
penguins |> 
  head(20) |> 
  write_csv("penguins20.csv")
```

```{r}
penguins_20 <- read_csv("penguins20.csv")

penguins_20
```

## AI use disclosure

(In a few sentences describe how you used AI in this lab.)

In this lab I used Github Copilot to help me to write this AI use disclosure.
